---
title: "Impact of Transmission on Fuel Efficiency"
author: "Eduardo Bortoluzzi Junior"
date: "November 15, 2015"
output: pdf_document
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, 
                      results='hide', fig.show='hide')
```

<!--
You work for Motor Trend, a magazine about the automobile industry. Looking at a 
data set of a collection of cars, they are interested in exploring the 
relationship between a set of variables and miles per gallon (MPG) (outcome). 
They are particularly interested in the following two questions:

* "Is an automatic or manual transmission better for MPG"
* "Quantify the MPG difference between automatic and manual transmissions"

Take the mtcars data set and write up an analysis to answer their question using 
regression models and exploratory data analyses.

* Did the student interpret the coefficients correctly?
* Did the student do some exploratory data analyses?
* Did the student fit multiple models and detail their strategy for model 
  selection?
* Did the student answer the questions of interest or detail why the question(s) 
  is (are) not answerable?
* Did the student do a residual plot and some diagnostics?
* Did the student quantify the uncertainty in their conclusions and/or perform 
  an inference correctly?
* Was the report brief (about 2 pages long) for the main body of the report and 
  no longer than 5 with supporting appendix of figures?
* Did the report include an executive summary?
* Was the report done in Rmd (knitr)?

-->

# Summary

This report describes the fuel consumption change between automobiles with manual
and automatic transmissions. This report was produced with automobile 
information extracted from the 1974 Motor Trend US magazine. As it is seen in 
the analysis, the fuel consumption is highly correlated to the weight of the
automobile and, in a lower level, to the gross horsepower, than any other
variable, so it was not conclusive if manual transmission cars has higher mpg
than the automatic ones.

The source code for the report is available at 
[this link][source].

[source]: https://github.com/sombrabr/regmods_courseproject/blob/master/transmissionmpg.Rmd

# Analysis

```{r}
require(ggplot2)
```

```{r}
# Multiple plot function
#
# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
# - cols:   Number of columns in layout
# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
#
# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
# then plot 1 will go in the upper left, 2 will go in the upper right, and
# 3 will go all the way across the bottom.
#
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```


```{r}
data(mtcars)
str(mtcars)
```

The data set *mtcars* has information on fuel consumption and 10 aspects of
automobile design and performance for 32 automobiles models of 1973 and 1974.

From physics, force (F) is defined as $F = m * a$, where *m* is mass and *a* is
acceleration; work (W) as $W = F * s$, where *s* is the displacement of the
object and energy (E) is needed to produce the work
($\partial W = \partial Q - dE$). Some heat (Q) is also produced in this
process.

For the internal combustion engine, the mechanical energy is produced by the
combustion of the fuel. So, the amount of combustion needed to make a automobile
displacement will be highly dependent on the automobile mass.

The plot1 has some observations related to the mpg. By the plot, it seems that
manual transmission has higher mpg than automatic one, but it does not correlate
other variables.

Using nested likelihood ratio tests, the addition of the variables about the 
engine construction will be tested on the hypothesis that the addition is not
significant to the anterior model.

```{r results='hold'}
fit0 <- lm(mpg ~ 1, mtcars)
fit1 <- update(fit0, ~ . + wt)
fit2 <- update(fit1, ~ . + hp)
fit3 <- update(fit2, ~ . + disp)
fit4 <- update(fit3, ~ . + cyl)

anova(fit0, fit1, fit2, fit3, fit4)
```

Only the addition of weight (`wt`) and gross horsepower (`hp`) are significant 
for the analysis, so the model `lm(mpg ~ wt + hp, mtcars)` is used in this
analysis. The plot2 shows the residuals of this model, and there are some
outliers that is influencing the model.

```{r}
fit <- lm(mpg ~ wt + hp, mtcars)

PRESS <- data.frame(press=resid(fit1) / (1 - hatvalues(fit1)))
rownames(PRESS) <- rownames(mtcars)

PRESS.mean <- mean(PRESS$press)
PRESS.sd <- sd(PRESS$press)

boundaries <- qnorm(c(0.025, 0.975), mean=PRESS.mean, sd=PRESS.sd)

sub <- subset(PRESS, PRESS$press < boundaries[1] | PRESS$press > boundaries[2])

mtcars.n <- mtcars[!(rownames(mtcars) %in% rownames(sub)),]

fit.n <- lm(mpg ~ wt + hp, mtcars.n)
```

Removing the observations which PRESS residuals is out of the 95% of the normal
density, which are `r rownames(sub)`, the new residuals will be those in plot3.

Now, adding the transmission information (*am*) to this model, the influence of
the type of the transmission is below:

```{r results='hold'}
fit.am <- update(fit.n, ~ . + factor(am))
summ.am <- summary(fit.am)
confint.am <- confint(fit.am)

summ.am
confint.am
```

* `r round(summ.am$r.squared * 100, 2)`% is the percentage of the total 
  variability that is explained by this model.
* `r round(summ.am$coefficients["factor(am)1","Estimate"], 2)`, varying from 
  `r round(confint.am["factor(am)1",1], 2)` to 
  `r round(confint.am["factor(am)1",2], 2)` defines the increase on mpg when 
  manual transmission is considered, and it is seen that
  it can increase or decrease the mpg in the confidence interval. Also, as the
  probability of getting values as extreme as 
  `r round(summ.am$coefficients["factor(am)1","t value"], 2)` is high 
  (`r round(summ.am$coefficients["factor(am)1","Pr(>|t|)"] * 100, 2)`%), we fail 
  to reject the hypothesis of automatic transmission having less full 
  consumption.
* As the weight variable is very significant, it is seen that mpg is more
  correlated to this variable.
  
# Conclusions

It is not conclusive whether manual or automatic has better MPG. The analysis
shows that the weight has higher correlation to the MPG.
  

# Appendix

## plot1
```{r plot1, fig.show='asis', fig.height=6}
g1 <- ggplot(mtcars, aes(wt, mpg)) +
      geom_point(alpha=.5) +
      geom_smooth(method="lm") +
      ggtitle("mpg X wt")

g2 <- ggplot(mtcars, aes(hp, mpg)) +
      geom_point(alpha=.5) +
      geom_smooth(method="lm") +
      ggtitle("mpg X hp")

g3 <- ggplot(mtcars, aes(disp, mpg)) +
      geom_point(alpha=.5) +
      geom_smooth(method="lm") +
      ggtitle("mpg X disp")

g4 <- ggplot(mtcars, aes(factor(am), mpg)) +
      geom_boxplot() +
      scale_x_discrete(labels=c("Automatic", "Manual")) +
      ggtitle("mpg X am")

multiplot(g1, g2, g3, g4, cols=2)
```


## plot2
```{r plot2, fig.show='asis'}
plot(fit, which=1)
```

## plot3
```{r plot3, fig.show='asis'}
plot(fit.n, which=1)
```
